EEG Waveform Analysis of P300 ERP with applications to Brain Computer Interfaces

### 2018-10-25 BrainScience/response.letter.1.tex

6. The Results section is not well-presented. The authors are advised to include a paragraph before Table 2, Table 3 and Fig. 6-8 presenting the obtained results (best/worst channel performance and algorithm performance for each experiment) and then discuss the results in the Discussion section. Please revise excessive descriptions in the captions of Table 2, Table 3 and Fig. 6-8.

7. “the obtained character identification rate is above theoretical chance level, and for some algorithms close to the usable threshold of 70\%” (Page 19, Paragraph 2, Line 465-466) What is the obtained character identification rate in other scientific studies? Is the obtained result (70\%) of this study comparable to others?

8. In the Discussion section a table presenting the statistical analysis results would present better the results and would facilitate the reader.

\subsection*{\ovalbox{General Comments}}
The paper is an interesting research paper and presents an analysis of the P300 ERP. The authors provide a literature review of the current methods used to detect EEG patterns and conduct three experiments to identify the P300 ERP in EEG recordings. The paper is well-organized. The paper should incorporate the following comments before its acceptance for publication:
\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for your comments and for this high quality and detailed review.  We have modified several points according to your recommendations. We hope the article is now better suited for publication. 
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

1. The major concern is that the experiments have been conducted with only one participant. The authors should collect more data in this study. In other studies cited in the paper by authors,  two and 16 participants were recruited.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
We have now included more participants to this experiment.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

2. It is strongly recommended that the authors should re-check some paragraphs of the manuscript, correct any grammatical and syntactic errors and improve the scientific language. In some cases, it was hard to follow the meaning. (e.g. Page 2, Paragraph 4, Line 51-54, Page 11, Paragraph 3, Line 343-345)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you very much for your comment.  We have modified the specific paragraphs that the Reviewer kindly offered as example and we have revised the entire manuscript modifying different parts, simplifying sections.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

### 2018-10-25 BrainScience/brainscience.tex
Subo un monton de cambios previos a poner todo otra vez

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{The Electroencephalography (EEG) is not just a mere clinical tool anymore.  It has become the de-facto mobile, portable, non-invasive brain imaging sensor to harness brain information in real time.  It is now being used to translate or decode brain signals to diagnose disease or implement Human Computer Interaction devices.  The automatic processing approach which is based on using quantitative algorithms to detect the cloaked information buried in the signal, overshadow the research done by the clinical EEG community.  Clinical EEG is based intensively on waveforms and the structure of signal plots.  Hence, the purpose of this work is to establish a bridge to fill this gap by reviewing and describing the procedures that have been used to detect patterns in the electroencephalographic waveforms, benchmarking them on a controlled pseudo-real dataset of an EEG Event Related Potentials waveform and verifying their performance on a public dataset of a BCI Competition.}

@@

EEG sensors are wearable~\citep{Puce2017} non-invasive, portable and mobile~\citep{DeVos2014}, with excellent temporal resolution, and acceptable spatial resolution~\citep{Hartman2005}.  This humble diagnosis device is been transformed into currently the best approach to detect, out-of-the lab in an ambulatory context, information from the Central Nervous System and to use that information to volitionally drive cars, steer drones, write emails, or control wheelchairs~\citep{Yuste2017}.

@@

Important applications have been developed in Spelling areas like this and that, in passive BCI with alcohol consumption monitoring.

KevricJasmin2016
Tzimourta2018
Cruz2018

Measure the flickering effect on the original vs the modified signal.

@@

Sleep Research has been studied in this way by performing Polysomnographic recordings (PSG)  \citep{Rodenbeck2006}, where the different sleep stages are evaluated by visually marking waveforms or graphoelements in long-running electroencephalographic recordings, looking for patterns based on standardized guidelines.   Visual characterization includes the identification or classification of certain waveform components, or transient events, based on a subjective characterization (e.g. positive or negative peak polarity) or the location within the strip.  It is regular to establish an amplitude difference between different waveforms from which a relation between them is reckoned and a structured index is created (e.g. sleep K-Complex is well characterized based on rates between positive vs negative amplitude) \citep{Uchida1999}.  Other relevant EEG patterns for sleep stage scoring are alpha, theta, and delta waves,  sleep spindles, polysplindles, vertex sharp waves (VSW), and sawtooth waves (REM Sleep).

@@

Regarding results produced for the public and real dataset IIb of P300 ERP from the Berlin BCI Competition II (2003), the obtained character identification rate is above theoretical chance level, and for some algorithms close to the usable threshold of $70\%$~\citep{Kathner2017}.  It is also similar to the performance obtained for the Experiment 3, which represents coincidentally the more realistic scenario for the pseudo-real dataset.  It is important to remark that the algorithms presented here analyze the waveform structure of a single-channel signal~\citep{Gribonval2008}.

>>>>

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{The Electroencephalography (EEG) is not just a mere clinical tool anymore.  It has become the de-facto mobile, portable, non-invasive brain imaging sensor to harness brain information in real time.  It is now being used to translate or decode brain signals to diagnose disease or implement Human Computer Interaction devices.  The automatic processing approach which is based on using quantitative algorithms to detect the cloaked information buried in the signal, neglected the research done by the clinical EEG community.  Clinical EEG is based intensively on waveforms and the structure of signal plots.  Hence, the purpose of this work is to establish a bridge to fill this gap by reviewing and describing the procedures that have been used to detect patterns in the electroencephalographic waveforms, benchmarking them on a controlled pseudo-real dataset of an EEG Event Related Potentials waveform and verifying their performance on a public dataset of a BCI Competition.}

@@

EEG sensors are wearable~\citep{Puce2017} non-invasive, portable and mobile~\citep{DeVos2014}, with excellent temporal resolution, and acceptable spatial resolution~\citep{Hartman2005}.  This humble diagnosis device is been transformed into currently the best approach to detect, out-of-the lab in an ambulatory context, information from the Central Nervous System and to use that information to volitionally drive cars, steer drones, write emails, or control wheelchairs, alcohol consumption monitoring~\citep{Yuste2017,KevricJasmin2016,Tzimourta2018,Cruz2018}.

@@

@@

Sleep Research has been studied in this way by performing Polysomnographic recordings (PSG)~\citep{Boostani2017,Rodenbeck2006}, where the different sleep stages are evaluated by visually marking waveforms or graphoelements in long-running electroencephalographic recordings, looking for patterns based on standardized guidelines~\citep{Dimitriadis2018}. Visual characterization includes the identification or classification of certain waveform components, or transient events, based on a subjective characterization (e.g. positive or negative peak polarity) or the location within the strip.  It is regular to establish an amplitude difference between different waveforms from which a relation between them is reckoned and a structured index is created (e.g. sleep K-Complex is well characterized based on rates between positive vs negative amplitude) \citep{Uchida1999}.  Other relevant EEG patterns for sleep stage scoring are alpha, theta, and delta waves,  sleep spindles, polysplindles, vertex sharp waves (VSW), and sawtooth waves (REM Sleep).

@@

Regarding results produced for the public and real dataset IIb of P300 ERP from the Berlin BCI Competition II (2003), the obtained character identification rate is above theoretical chance level, and for some algorithms close to the usable threshold of $70\%$~\citep{Kathner2017}.  Results for this competition show perfect classification with tailored algorithm.  Even comparing the results with SVM multichannel, which uses all the available information from all channels show comparable results.

It is also similar to the performance obtained for the Experiment 3, which represents coincidentally the more realistic scenario for the pseudo-real dataset.  It is important to remark that the algorithms presented here analyze the waveform structure of a single-channel signal~\citep{Rakotomamonjy2008,Gribonval2008}.
